---
title: "lingtypology:"
subtitle: "easy mapping for linguistic typology in R"
author: "G. Moroz"
date: "ConCorT (15–17 October 2017) <br> HSE Voronovo Learning Center"
output:
  xaringan::moon_reader:
    css: ["default", "my_theme.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message=FALSE,
                      fig.width=10.4,
                      fig.height=5)
```

class: inverse, center, middle

# Why lingtypology?


---
## Why lingtypology?

* create uniform access to data across publications
  * [Glottolog](http://glottolog.org/)
  * [Cross-Linguistic Linked Data project](http://clld.org/)
* easy map creation
* multiple visualisations for typological features
* API to typological databases

---
class: inverse, center, middle
# Installation

---
## Installation
Install from CRAN (better):
```{r, eval = FALSE}
install.packages(lingtypology)
```

Install from GitHub:
```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("ropensci/lingtypology")
```

Load package:
```{r}
library(lingtypology)
```

In this presentation I use the following version:
```{r}
packageVersion("lingtypology")
```

---
class: inverse, center, middle
# Glottolog functions

---
## Glottolog functions
Structure of functions: **what you need - what you have**

---
## Glottolog functions: by language
```{r}
aff.lang("Adyghe")
area.lang(c("Basque", "Ewe", "Dyirbal", "Tobo"))
gltc.lang(c("Au", "Bau", "Iau", "Lau"))
iso.lang(c("Wan", "Han", "Dan"))
```

---
## Glottolog functions: get languages
```{r}
lang.aff("Kartvelian")
lang.iso(c("wan", "haa", "daf"))
lang.gltc(c("auuu1241", "bauu1244", "iauu1242", "lauu1247"))
```

---
## Glottolog functions: ISO 639-3 ↔ Glottocodes
```{r}
iso.gltc(c("auuu1241", "bauu1244", "iauu1242", "lauu1247"))
gltc.iso(c("wan", "haa", "daf"))
```

--

It is possible to use the output of one function as the input for another:
```{r}
gltc.lang(lang.aff("Circassian"))
```

---
## Glottolog functions: Abbreviations
```{r}
lang.country("Cape Verde")
lang.country("Cabo Verde")
head(lang.country("USA"))
```

---
## Glottolog functions: Spell Checker
```{r}
iso.gltc(c("auuu1241  ", " lau  u1247"))
aff.lang(c("Adyge", "Katalan"))
```

---
class: inverse, center, middle
# Creating a map

---
## Map: base map

```{r}
map.feature("Estonian")
```

---
## Map: multiple languages

```{r}
map.feature(lang.aff("Turkic"))
```

---
## Built-in datasets

```{r}
head(ejective_and_n_consonants)
head(circassian)
```

---
## Color languages by a categorical variable
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives)
```

---
## Color languages by a numeric variable
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$consonants)
```

---
## Change colors
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives,
            color = c("yellow", "navy"))
```

---
## Change colors
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$consonants,
            color = "magma")
```

---
## Add labels
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives,
            label = ejective_and_n_consonants$language)
```

---
## Add labels: permanent labels
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives,
            label = ejective_and_n_consonants$language,
            label.hide = FALSE)
```

---
## Add labels: emphasize
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives,
            label = ejective_and_n_consonants$language,
            label.hide = FALSE,
            label.emphasize = list(17:19, "red"))
```

---
## Add minicharts: bar
```{r}
map.feature(ejective_and_n_consonants$language,
            minichart.data = ejective_and_n_consonants[, 3:4],
            minichart = "bar")
```

---
## Add minicharts: pie
```{r}
map.feature(ejective_and_n_consonants$language,
            minichart.data = ejective_and_n_consonants[, 3:4],
            minichart = "pie")
```

---
## Add minicharts: pie
```{r}
map.feature(ejective_and_n_consonants$language,
            minichart.data = ejective_and_n_consonants[, 3:4],
            minichart = "pie",
            minichart.labels = TRUE)
```

---
## Add your own coordinates
```{r}
map.feature(circassian$language, features = circassian$dialect,
            latitude = circassian$latitude, 
            longitude = circassian$longitude)
```

---
## Add strokes
```{r}
map.feature(circassian$language, features = circassian$dialect,
            latitude = circassian$latitude, 
            longitude = circassian$longitude,
            stroke.features = circassian$language)
```

---
## Add a rectangle
```{r}
map.feature(circassian$language, features = circassian$dialect,
            latitude = circassian$latitude, 
            longitude = circassian$longitude,
            rectangle.lng = c(42.7, 45),
            rectangle.lat = c(42.7, 44.3))
```

---
## Add a density estimation
```{r}
map.feature(circassian$language, features = circassian$dialect,
            latitude = circassian$latitude, 
            longitude = circassian$longitude,
            density.estimation = circassian$language)
```

---
## Add a density estimation without dots
```{r}
map.feature(circassian$language, features = circassian$dialect,
            latitude = circassian$latitude, 
            longitude = circassian$longitude,
            density.estimation = circassian$language,
            density.points = FALSE)
```

---
## Change tiles
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives,
            tile = "OpenStreetMap.BlackAndWhite")
```

---
## Add minimaps
```{r}
map.feature(ejective_and_n_consonants$language,
            features = ejective_and_n_consonants$ejectives,
            minimap = TRUE)
```

---
class: inverse, center, middle
# Databases API

---
## Databases API

* `wals.feature()` World Atlas of Language Structures
* `autotyp.feature()` AUTOTYP
* `phoible.feature()` PHOIBLE
* `afbo.feature()`  Affix Borrowing database
* `sails.feature()`  South American Indigenous Language Structures
* `abvd.feature()` Austronesian Basic Vocabulary Database

---
## WALS

```{r, eval=FALSE}
df <- wals.feature("1a")
head(df)
```
```{r, echo=FALSE, warning = FALSE}
df <- wals.feature("1a")
df <- df[,-4]
head(df)
```


---
## AUTOTYP

```{r}
df <- autotyp.feature(c('Numeral classifiers'))
head(df)
```

---
class: inverse, center, middle
# About package development

---
## About package development

.pull-left[
* I've created
  * 13 issues in [ropensci/lingtypology](https://github.com/ropensci/lingtypology)
  * 4 issues in [clld/glottolog-data](https://github.com/clld/glottolog-data)
  * 1 issue in [clld/glottolog](https://github.com/clld/glottolog)
  * 1 issue in [clld/wals3](https://github.com/clld/wals3)
  * 1 issue in [clld/sails](https://github.com/clld/sails)
  * 1 issue in [rstudio/leaflet](https://github.com/rstudio/leaflet)
  * 1 issue in [yihui/xaringan](https://github.com/yihui/xaringan)
]

--

.pull-right[
* I received some comments and suggestions from
  * [Samira Verhees](https://github.com/sverhees)
  * [Calle Börstell](https://github.com/borstell)
  * [Robert Forkel](https://github.com/xrotwang)
  * Natalia Levshina
  * Timo Roettger
* and from [rOpenSci](https://github.com/ropensci) peer review
  * [Scott Chamberlain](https://github.com/sckott)
  * [Kent Russell](https://github.com/timelyportfolio/)
  * [Taras Tzakharko](https://github.com/tzakharko)
  * [Languagespacelabs](https://github.com/languagespacelabs)
]
---
class: middle

* Don't hesitate to write (<agricolamz@gmail.com>) or open [an issue](https://github.com/ropensci/lingtypology/issues).

* `lingtypology` tutorial: <https://ropensci.github.io/lingtypology/>

* Slides on **GitHub**: <https://agricolamz.github.io/2017_ConCorT_lingtypology>


* Slides created with the beautiful R package [**xaringan**](https://github.com/yihui/xaringan)
